Bayesian Deep Gaussian Processes for Correlated Functional Data: \\ A Case Study in Cosmological Power Spectra

### 2025-06-26 paper/main.tex
edits to 4.2

In order to predict power spectra from the CAMB dataset, we split the 64 cosmologies such that $m=32$
are used to train the DGP.FCO model and the remaining $N=32$ are reserved as hold-out cosmologies for
prediction. With our $m=32$ posterior means using DGP.FCO, we obtain $N=32$ predicted power spectra from
the functional prediction model outlined in Section \ref{subsec:pca} for the test cosmologies.

\begin{figure}
    \centering
    \includegraphics[width=0.85\textwidth]{pred_diffs_CAMB.jpeg}
    \caption{Predictions of power matter spectra for each of the 32 CAMB test cosmologies,
             after centering and scaling (to accentuate differences). 
             The true infinite resolution run is shown in black, the prediction 
             trained on the infinite resolution spectra is in dashed blue, and the prediction from 
             the DGP.FCO posterior means is in dotted orange.}
    \label{fig:pca_preds_v_camb}
\end{figure}

Since CAMB offers the ``true'' infinite-resolution spectra, we have the opportunity to benchmark our
results in two ways.  First, we may compare our predictions to the true power spectrum for each held-out
cosmology.  Second, we may leverage the infinite-resolution spectra for the training cosmologies in
place of our estimated $\hat{S}_j$, to get a sense of the ``best-case" predictions that could originate
from the PC decomposition.  Figure \ref{fig:pca_preds_v_camb} compares our predictions (DGP.FCO + PC, orange)
to these ``best-case'' predictions (true + PC, blue) and the actual true spectra for each of the 32
held-out test cosmologies.  There is a high level of agreement between the pairs of predictions across 
cosmologies. 
We then compare the mean squared errors of our predictions and the ``best-case'' 
predictions across all held-out cosmologies, with results shown in Figure \ref{fig:mse_camb}.  We again
see only minimal differences, which we take to indicate that DGP.FCO is offering effective
estimates of the true spectra for each cosmology.

>>>>

For the CAMB dataset, we split the 64 available cosmologies such that $m=32$
are used for training and $N=32$ are reserved as hold-out/test cosmologies for prediction. 
For each training cosmology, we collect our DGP.FCO predicted spectrum, then conduct the principal
components model just described to predict $\hat{S}(\psi)$ for each test cosmology.

\begin{figure}
    \centering
    \includegraphics[width=0.85\textwidth]{pred_diffs_CAMB.jpeg}
    \caption{Predictions of power matter spectra for each of the 32 CAMB test cosmologies,
             after centering and scaling (to accentuate differences). 
             The true infinite resolution run is shown in black, the PC prediction 
             trained on the infinite resolution spectra is in dashed blue, and the PC prediction trained
             on the DGP.FCO posterior means is in dotted orange.}
    \label{fig:pca_preds_v_camb}
\end{figure}

Since CAMB offers the ``true'' infinite-resolution spectra, we have the opportunity to benchmark our
results in two ways.  First, we may compare our predictions to the true power spectrum for each held-out
cosmology.  Second, we may leverage the infinite-resolution spectra for the training cosmologies in
place of our estimated $\hat{S}_j$ from our DGP.FCO model, to get a sense of the ``best-case" predictions 
that could originate from the PC decomposition.  
Figure \ref{fig:pca_preds_v_camb} compares our predictions (DGP.FCO + PC, dotted orange)
to these ``best-case'' predictions (true + PC, dashed blue) and the actual true spectra (solid black)
for each of the 32 held-out test cosmologies.  
The agreement between our predictions and the ``best-case'' predictions indicates that DGP.FCO is effectively
capturing the true power matter spectra of the training cosmologies.
The high level of agreement between the pairs of predictions and the true spectra across cosmologies
indicates our PC model is providing effective predictions of held-out power spectra.
We then compare the mean squared errors of our predictions and the ``best-case'' 
predictions across all held-out cosmologies, with results shown in Figure \ref{fig:mse_camb}.  We again
see only minimal differences between the two, confirming that DGP.FCO is offering effective
estimates of the true spectra for each cosmology.
